### 2021-05-13 assertions/doc/thesis-bernhardt/chapters/3-offline_design.tex
Final revision part 1

\chapter{Toolchain Implementation - Frontend}\label{chap:offline}
The purpose of the toolchain is to support the process from specifying assertions to generating assertion Smart Contracts that can be invoked on-chain by the validators. As determined in \secref{sec:restrict}, the current version of the toolchain only supports the processing of assertions featuring universal quantification. The toolchain was developed in the programming language OCaml \cite{ocaml_docs} --- the code is available in a GitHub\footnote{\url{https://github.com/proglang/thesis-distributed-assertion-checking}\todo{commit}} repository of the Chair of Programming Languages.

@@

\section{Assertion syntax}\label{sec:syntax}
Tezos' Smart Contract language Michelson is a low-level and stack-based language. Specifying assertions in such a target-specific language is complex and inconvenient, especially if the language should be compiled to other targets as well. Therefore, this thesis defines a domain-specific high-level language to express assertions.

@@

\subsection{Code structure}
Assertions are passed to the toolchain as simple text files. This thesis follows the convention of using the file extension \texttt{.tza} for such files (henceforth referred to as ``assertion contracts'') in reference to the file extension \texttt{.tz} used for Michelson source files. An assertion contract is always associated with a Smart Contract (henceforth referred to as ``parent contract'') and contains at least one assertion for a specific entrypoint thereof. For the remaining entrypoints, the compiler automatically generates empty assertions (cf. \secref{sec:compiler}).

@@

How the assertions are associated with the respective entrypoint of the parent contract, depends on the Smart Contract language. In the case of Solidity, a tag that mirrors the function name is enough to link assertion and entrypoint. In Michelson contracts, on the other hand, entrypoints can be anonymous. Therefore, the assertions are linked to the entrypoints through the tags, the parameter type or both. \secref{sec:typecheck} describes in more detail how the linking is implemented in the Tezos backend.

\subsection{Parameter patterns}
\lstref{lst:patterns} gives an overview of the parameter patterns supported in the assertion language.
\lstinputlisting[caption=Patterns in the assertion language, language=, numbers=none, label=lst:patterns]{listings/patterns.txt}
By using patterns, developers can avoid the explicit handling of special cases, such as empty lists, in the assertion specification. Furthermore, they can be used to put emphasis on the relevant components of the input parameters by replacing the irrelevant ones with wildcards. As an example, the assertion given in \lstref{lst:patterns_usage} concerns the left element of a pair and only applies to non-empty lists. 
\begin{lstlisting}[label=lst:patterns_usage, language=Assertion, numbers=none]
(entrypoint %g (pair (cons a as) _)  ...)
\end{lstlisting}

\subsection{Types and operations}
In order to make the frontend generic, the sets of data types and operations supported by the grammar constitute unions of those present in the target languages. This allows not only the implementation of a single, common parser, but also the transformation to be oblivious of the target platform. As a consequence, the backends are responsible for performing a semantic check and  rejecting any assertions containing types or operations which are not supported by the target language.

@@

The grammar currently includes common arithmetic (e.g., \texttt{add}, \texttt{sub}, \texttt{abs}, \texttt{mod}), relational  (e.g., \texttt{eq}, \texttt{neq}, \texttt{lt}) and logic and bitwise operators (e.g., \texttt{and}, \texttt{or}, \texttt{lsr}, \texttt{lsl}). Some operations that operate on data structures (e.g., \texttt{nth} and \texttt{size}) are supported as well. Domain specific operations have not been included yet, but they may be added to the language if necessary.

\subsection{Example}
As an example, consider a variation of the explicit assertion for sorted lists, stated in \eqref{eq:sorted_v2}. In contrast to the variation given in the introduction (\eqref{eq:sorted}), which translates to comparing a random element of the list with its direct neighbour, this translates to a comparison of two random elements (under the constraint $n < m$). \lstref{lst:sorted_assertion} shows how this formula can be expressed in the assertion language.
\begin{equation}\label{eq:sorted_v2}
	(\forall n,m : int) (0 \leq n < m < |a|) \Rightarrow a[n] \leq a[m]
\end{equation}
\lstinputlisting[caption=Assertion specification for sorted lists, language=Assertion, label={lst:sorted_assertion}]{listings/sorted.tza}

Line 1 declares the signature of the assertion. It specifies the explicit tag \textit{find} and the input parameter \texttt{a} of type integer list. The body is a nesting of two universal quantifiers, which declare the predicate variables $n$ and $m$ of type integer (lines 2 and 3), followed by a nesting of conditions. The conditions define the domains of the quantifiers (lines 4-6), corresponding to the premise of the formula. The body of the last condition (line 7) ultimately contains the keyword \texttt{assert} followed by a condition checking the property of ascending order, i.e.,  $a[n] \leq a[m]$.

\subsection{Extensions}\label{sec:ext}
In future iterations, the assertion syntax could be extended with some more features to improve usability and readability:
\begin{itemize}
\item \textbf{Local variables} to store, reuse and denominate computed values
\item \textbf{User-defined functions} to extract whole routines, s.t. they can be reused in several assertions or other assertion contracts
\item \textbf{if-else conditions} to adapt the domains of quantifications if certain conditions hold. The assertion checking whether two numbers are relatively prime (discussed in \secref{sec:coprime}) is a good use-case for this feature: Firstly, the formula requires to determine the minimum of the given numbers --- however, the \texttt{min}-function might not be supported as a built-in function by the target language (as is the case on Tezos). A solution for that is to implement two branches in the program to handle each case. Secondly, if the greater of the two numbers is not evenly divided by the smaller one, the formula can be optimized by reducing the quantification domain by half to $2 \le n \le \lfloor \frac{min(a,b)}{2} \rfloor$. \lstref{lst:coprime} shows a hypothetical specification of an assertion validating the property of coprimeness using \texttt{if-else} control structures:
\lstinputlisting[caption=Assertion for coprimeness using if-else structures, language=Assertion, label={lst:coprime}]{listings/coprime_ifelse.tza}
\end{itemize}

@@

\figref{fig:ast_example} shows the resulting AST after parsing the assertion given in \lstref{lst:sorted_assertion}. The subtype representing quantifiers contains a special field \textit{Bounds}, which is initially left empty by the parser. During the transformation, it is populated with a set of constraints concerning its predicate variable, cf. \secref{sec:transformation}.

\section{Transformation}\label{sec:transformation}
The assertion, as specified by a developer, is modified in two ways before the compiler translates it into target code: firstly, it has to be transformed into its negation. Secondly, quantifiers are converted to random generators, which generate values within a specified range (corresponding to the domains of discourse). In the assertion language, the ranges can be expressed using \texttt{if}-conditions. As a preliminary for the compilation, the transformation additionally analyses the condition nodes of the AST, identifies conditions that state a range constraint for a generator and assigns it to the respective quantification node (by adding it to the \textit{Bounds} field). Therefore, the transformation contains two consecutive steps:
\begin{enumerate}
\itemsep-0.5em
\item Negation
\item Domain constraint analysis and assignment
\end{enumerate}
Each step is described in more detail in the following subsections.

@@

\subsection{Domain assignment}
In order to obtain random generators which only generate values in the interesting range, i.e., within the domains of discourse, the second step of the transformation skims the condition nodes of the AST for atomic constraints containing bound variables. These conditions are then moved to the \texttt{Bounds} field of the respective quantifier binding that variable. If the constraint contains more than one bound variable, it is assigned to the quantifier with the highest depth in the quantifier nesting. When translating the quantifiers to random generators, their ranges can then be constructed based on the associated list of constraints. After merging a constraint with a quantifier node, it is removed from the assertion body, as checking these constraints during runtime becomes redundant. Constraints not involving bound variables have to be checked during runtime.

Relating to the example of sorted lists, the premise in \eqref{eq:sorted_v2} can be considered as a conjunction of the four constraints
\begin{enumerate}
\itemsep-1em
\item $0 \leq n$
\item $n < |a| $
\item $n \le m$ and
\item $m < |a|$.
\end{enumerate}
Constraint 1,2 and 4 only contain one bound variable ($n$ or $m$), thus they are assigned to the respective quantifier binding it. Constraint 3, on the other hand, features two bound variables -- this constraint is assigned to the quantifier furthest in the nesting, i.e., the quantifier binding $m$. The corresponding AST expressing this assignment is shown in \figref{fig:ast_example_transf}. Based on the assigned constraints, the quantifiers can then be translated to target code corresponding to the following pseudocode (\lstref{lst:rand}):
\begin{lstlisting}[label=lst:rand, caption=Random generators for sorted list assertion, numbers=none]
n = random(0, size(a))
m = random(n + 1, size(a))
\end{lstlisting}

However, if constraints contain certain operators, deriving efficient random generators is less straight-forward. Consider the following constraints:
\begin{enumerate}
\itemsep-0.7em
\item $(\forall n : int) (n < 10 \lor n > 20) ...$
\item $(\forall n : int) (n \ne 10) ...$
\item $(\forall n : int) (n = 10) ...$
\end{enumerate}
Constraint 1 can complicate restricting the range of the random generator in two ways. Firstly, it separates the range into two disjoint sub-ranges, which translates to a random generator as shown in \lstref{lst:rand_disjoint}.
\begin{lstlisting}[label=lst:rand_disjoint, caption=Random generator with two disjoint ranges, numbers=none]
n = if random(0,2) == 0 then random(-MAX_INT, 10)
                        else random(21, MAX_INT)
\end{lstlisting}
Secondly, the operands of a disjunction cannot be considered separately if they involve several bound variables. For the predicate variable whose random generator is executed first, the domain restriction is optional, while the generation of the following random values depends on the previous result. \lstref{lst:rand_disjoint_nm} shows the corresponding random generators for a variation of constraint 1 featuring two bound variables, i.e., $(\forall n,m) (n < 10 \lor m > 20)...$.
\begin{lstlisting}[label=lst:rand_disjoint_nm, caption=Random generators of a disjunction of constraints on two bound variables, numbers=none]
n = if random(0,2) == 0 then random(-MAX_INT, 10)
                        else random(-MAX_INT, MAX_INT)
m = if n < 10 then random(-MAX_INT, MAX_INT)
              else random(20, MAX_INT)
\end{lstlisting}
Similarly to 1, constraint 2 splits the domain into sub-domains. Constraint 3 effectively makes the quantifier binding $n$ obsolete and can therefore be translated to a variable with a constant value.

In the current implementation of the transformation, constraints featuring the logical or, exclusive or or inequality operator are kept as part of the assertion body. They are considered to be exceptional cases, but will increase the overall complexity of the compilation beyond the scope of a first iteration. In the case that constraints remain part of the assertion body, they are checked during runtime. As a consequence, an invocation of the assertion contract can result in a futile assertion check if such a condition is not met.

\subsubsection{Implicit constraints}
In some cases, boundaries are imposed to the random generators implicitly through the used data types. List indices, for instance, are always bound to the range $[0.. size(list))$ and could thus be derived from the formula without an explicit specification. As this requires a semantic analysis of the AST and thus a more complex transformation, the current implementation requires a completely explicit formula in terms of the domains.

>>>>

\chapter{Toolchain Frontend}\label{chap:offline}
The purpose of the toolchain is to support the process from specifying assertions to generating assertion smart contracts that can be invoked on-chain by the validators. As determined in \secref{sec:restrict}, the current version of the toolchain only supports the processing of assertions featuring universal quantification. The toolchain was developed in the programming language OCaml \cite{ocaml_docs} --- the code is available in a GitHub\footnote{\url{https://github.com/proglang/thesis-distributed-assertion-checking/commit/7c22038cebe44f779621d241a8ab156e105ead0c}} repository of the Chair of Programming Languages.

@@

\section{Assertion Syntax}\label{sec:syntax}
Tezos' smart contract language Michelson is a low-level and stack-based language. Specifying assertions in such a target-specific language is complex and inconvenient, especially if the language should be compiled to other targets as well. Therefore, this thesis defines a domain-specific high-level language to express assertions.

@@

\subsection{Code Structure}
Assertions are passed to the toolchain as simple text files. This thesis follows the convention of using the file extension \texttt{.tza} for such files (henceforth referred to as ``assertion contracts''), in reference to the file extension \texttt{.tz} used for Michelson source files. An assertion contract is always associated with a smart contract (henceforth referred to as ``parent contract'') and contains at least one assertion for a specific entrypoint thereof. For the remaining entrypoints, the compiler should automatically generate empty assertions (cf. \secref{sec:compiler}).

@@

How the assertions are associated with the respective entrypoint of the parent contract, may depend on the smart contract language. In the case of Solidity, a tag that mirrors the function name is enough to link assertion and entrypoint. In Michelson contracts, on the other hand, entrypoints can be anonymous. Therefore, the assertions are linked to the entrypoints through the tags, the parameter type or both. \secref{sec:typecheck} describes in more detail how the linking is implemented in the Tezos backend.

\subsection{Parameter Patterns}
\lstref{lst:patterns} gives an overview of the parameter patterns supported in the assertion language.
\lstinputlisting[caption=Patterns in the assertion language, language=, numbers=none, label=lst:patterns]{listings/patterns.txt}
By using patterns, developers can avoid the explicit handling of special cases, such as empty lists, in the assertion specification. Furthermore, they can be used to put emphasis on the relevant components of the input parameters by replacing the irrelevant ones with wildcards. As an example, the assertion given in \lstref{lst:patterns_usage} concerns the left element of a pair and only applies to non-empty lists. 
\begin{lstlisting}[label=lst:patterns_usage, language=Assertion, numbers=none]
(entrypoint %g (pair (cons a as) _)  ...)
\end{lstlisting}

\subsection{Types and Operations}
In order to make the frontend generic, the sets of data types and operations supported by the grammar constitute unions of those present in the target languages. This allows not only the implementation of a single, common parser, but also the transformation to be oblivious of the target platform. As a consequence, the backends are responsible for performing a semantic check and  rejecting any assertions containing types or operations which are not supported by the target language.

@@

The grammar currently includes common arithmetic (e.g., \texttt{add}, \texttt{sub}, \texttt{abs}, \texttt{mod}), relational  (e.g., \texttt{eq}, \texttt{neq}, \texttt{lt}) and logic and bitwise operators (e.g., \texttt{and}, \texttt{or}, \texttt{lsr}, \texttt{lsl}). Some operations that operate on data structures (e.g., \texttt{nth} and \texttt{size}) are supported as well. Domain specific operations have not been included yet, but they may be added to the language if necessary. \\

Let us consider a variation of the explicit assertion for sorted lists, stated in \eqref{eq:sorted_v2}. In contrast to the variation given in the introduction (\eqref{eq:sorted}), which translates to comparing a random element of the list with its direct neighbour, this translates to a comparison of two random elements (under the constraint $n < m$). \lstref{lst:sorted_assertion} shows how this formula can be expressed in the assertion language.
\begin{equation}\label{eq:sorted_v2}
	(\forall n,m : int) (0 \leq n < m < |a|) \Rightarrow a[n] \leq a[m]
\end{equation}
\lstinputlisting[caption=Assertion specification for sorted lists, language=Assertion, label={lst:sorted_assertion}]{listings/sorted.tza}

Line 1 declares the signature of the assertion. It specifies the explicit tag \textit{find} and the input parameter \texttt{a} of type integer list. The body is a nesting of two universal quantifiers, which declare the predicate variables $n$ and $m$ of type integer (lines 2 and 3), followed by a nesting of conditions. The conditions define the domains of the quantifiers (lines 4-7), corresponding to the premise of the formula. The body of the last condition (line 7) ultimately contains the keyword \texttt{assert} followed by a condition checking the property of ascending order, i.e.,  $a[n] \leq a[m]$.

\subsection{Extensions}\label{sec:ext}
In future iterations, the assertion syntax could be extended with some more features to improve usability and readability:
\begin{itemize}
\item \textbf{Local variables} to store, reuse and denominate computed values
\item \textbf{User-defined functions} to extract whole routines, s.t. they can be reused in several assertions or other assertion contracts
\item \textbf{if-else conditions} to adapt the domains of quantifications if certain conditions hold. The assertion checking whether two numbers are relatively prime (introduced in \secref{sec:coprime}) is a good use-case for this feature: firstly, the formula requires to determine the minimum of the given numbers --- however, the \texttt{min}-function might not be supported as a built-in function by the target language (as is the case in Michelson). A solution for that is to implement two branches in the program to handle each case. Secondly, if the greater of the two numbers is not evenly divided by the smaller one, the formula can be optimized by reducing the quantification domain by half to $2 \le n \le \lfloor \frac{min(a,b)}{2} \rfloor$. \lstref{lst:coprime} shows a hypothetical specification of an assertion validating the property of coprimeness using \texttt{if-else} control structures:
\lstinputlisting[caption=Assertion for coprimeness using if-else structures, language=Assertion, label={lst:coprime}]{listings/coprime_ifelse.tza}
\end{itemize}

@@

\figref{fig:ast_example} shows the resulting AST after parsing the assertion given in \lstref{lst:sorted_assertion}. The subtype representing quantifiers contains a special field \textit{Bounds}, which is initially left empty by the parser. During the transformation, it is populated with a set of constraints concerning its predicate variable. This is described in the following section.

\section{Transformation}\label{sec:transformation}
The assertion, as specified by a developer, is modified in two ways before the compiler translates it into target code: firstly, it has to be transformed into its negation. Secondly, quantifiers are converted to random generators, which generate values within a specified range (corresponding to the domains of discourse). In the assertion language, the ranges can be expressed using \texttt{if}-conditions. As a preliminary for the compilation, the transformation thus analyses the condition nodes of the AST, identifies conditions that state a range constraint for a generator and assigns it to the respective quantification node (by adding it to the \textit{Bounds} field). Therefore, the transformation contains two consecutive steps:
\begin{enumerate}
\itemsep-0.5em
\item Negation
\item Domain constraint analysis and assignment
\end{enumerate}
Each step is described in more detail in the following subsections.

@@

\subsection{Domain Assignment}\label{sec:transf_domains}
In order to obtain random generators which only generate values in the interesting range, i.e., within the domains of discourse, the second step of the transformation skims the condition nodes of the AST for atomic constraints containing bound variables. These conditions are then moved to the \texttt{Bounds} field of the respective quantifier binding that variable. If the constraint contains more than one bound variable, it is assigned to the quantifier with the highest depth in the quantifier nesting. When translating the quantifiers to random generators, their ranges can then be constructed based on the associated list of constraints. After merging a constraint with a quantifier node, it is removed from the assertion body, as checking these constraints during runtime becomes redundant. Constraints not involving bound variables have to be checked during runtime.

The assertion for sorted lists contains the four constraints
\begin{enumerate}
\itemsep-1em
\item $0 \leq n$
\item $n < |a| $
\item $n \le m$ and
\item $m < |a|$.
\end{enumerate}
Constraint 1,2 and 4 only contain one bound variable ($n$ or $m$), hence they are assigned to the respective quantifier binding it. Constraint 3 features two bound variables -- this constraint is assigned to the quantifier furthest in the nesting, i.e., the quantifier binding $m$. The corresponding AST expressing this assignment is shown in \figref{fig:ast_example_transf}. Based on the assigned constraints, the compiler can then translate the quantifiers to the random generators as indicated in \lstref{lst:rand}.
\begin{lstlisting}[label=lst:rand, caption=Random generators for the sorted list assertion, numbers=none]
n = random(0, size(a))
m = random(n + 1, size(a))
\end{lstlisting}

For constraints containing certain operators, deriving efficient random generators is less straight-forward. Consider the following constraints:
\begin{enumerate}
\itemsep-0.7em
\item $(\forall n : int) (n < 10 \lor n > 20) ...$
\item $(\forall n,m : int) (n < 10 \lor m > 20)$
\item $(\forall n : int) (n \ne 10) ...$
\item $(\forall n : int) (n = 10) ...$
\end{enumerate}
Constraint 1 separates the range into two disjoint sub-ranges, which translates to a random generator as shown in \lstref{lst:rand_disjoint}.
\begin{lstlisting}[label=lst:rand_disjoint, caption=Random generator with two disjoint ranges, numbers=none]
n = if random(0,2) == 0 then random(-MAX_INT, 10)
                        else random(21, MAX_INT)
\end{lstlisting}
If a disjunction involves several bound variables, as in constraint 2, the operands of this disjunction cannot be considered separately, as would be the case with a conjunction. For the predicate variable whose random generator is executed first, the domain restriction is optional, while the generation of the following random value depends on the previous result. \lstref{lst:rand_disjoint_nm} shows the corresponding random generators for this case.
\begin{lstlisting}[label=lst:rand_disjoint_nm, caption=Random generators of a disjunction of constraints with two bound variables, numbers=none]
n = if random(0,2) == 0 then random(-MAX_INT, 10)
                        else random(-MAX_INT, MAX_INT)
m = if n < 10 then random(-MAX_INT, MAX_INT)
              else random(20, MAX_INT)
\end{lstlisting}
Similarly to 1, constraint 3 splits the domain into sub-domains. Constraint 4 effectively makes the quantifier binding $n$ obsolete and can therefore be translated to a constant value.

In the current implementation of the transformation, constraints featuring the logical or, exclusive or inequality operator are kept as part of the assertion body. They are considered to be exceptional cases, but will increase the overall complexity of the compilation beyond the scope of a first iteration. Constraints remaining within the assertion body will be checked during runtime. As a consequence, an invocation of the assertion contract can result in a futile assertion check if such a condition is not met.

\subsubsection{Implicit Constraints}
In some cases, boundaries are imposed to the random generators implicitly by the used data types. List indices, for instance, are always bound to the range $[0.. size(list))$ and could thus be derived from the formula without an explicit specification. As this requires a semantic analysis of the AST and thus a more complex transformation, the current implementation requires a completely explicit formula in terms of the domains.

### 2021-05-13 assertions/doc/thesis-bernhardt/chapters/4-probabilistic_model.tex
Revision compiler
Final revision part 1

\chapter{Probabilistic model}\label{chap:prob_model}
Since the proposed approach implements a probabilistic validation, the reliability of the validation depends on the number of validators in the network in relation to the size of input. Given that the number of validators is finite, the probability of accepting an invalid result may exceed an acceptable threshold for big inputs. Thus, the distributed assertion checking scheme should provide a mechanism to have each validator invoke the assertion contract a specified number of times. To this end, it is necessary to identify a formula which, given the input size and a probability threshold, returns a lower bound for the required number of test runs per validator.

@@

\section{Sample spaces}
The sample space of an assertion is given by the domains of discourse of the quantifiers. \secref{sec:transf_domains} described, that the premises defining the domains are translated into a list of range constraints for the random generators. Based on these constraints, the compiler can derive a formula to determine the size of the sample space at runtime. In simple cases, the sample space is given by a single domain. Such a space is given, for instance, in the introductory example checking whether a number is a prime, i.e., $\mathcal{S} = \lbrace n \in\mathbb{Z} : 2 \le n \le \sqrt{p} \rbrace$. Based on the constraints $2 \le n$ and $n \le \sqrt{p}$, the size of the search space can be determined at runtime with $|\mathcal{S}| = \sqrt{p} - 2$. If the assertion has to sample from more than one domain, the search space is given by the Cartesian product of these domains. In the example of intersecting sets from \secref{sec:existential}, the sample space is $\mathcal{S} = \mathcal{S}_U \times \mathcal{S}_V$, where $\mathcal{S}_U = \lbrace n \in\mathbb{Z} : 0 \leq n \le |U| \rbrace$ and $\mathcal{S}_V = \lbrace m \in\mathbb{Z} : 0 \leq m \le |V| \rbrace$). At runtime, the size of the search space can be determined with $|\mathcal{S}| = (|U| - 0) + (|V| - 0)$.

@@

\section{Configuration of the error rate}\label{sec:prob_threshold}
The goal is to find a number $t$ of test runs, s.t. the probability $P_{c,t}$ of accepting an invalid result falls below a certain threshold $c$. A validator finds the counterexample in one test run with probability $\frac{1}{n}$, and misses it with probability $(1-\frac{1}{n})$. After $t$ tests runs, the probability that the counterexample has not been found is thus $(1-\frac{1}{n})^t$. Following the same approach as Mahlmann and Schindelhauer \cite{mahl_schindel_2007} to retrieve an upper bound for $P_{c,t}$, the following inequality is used:
\begin{equation}
(1-\frac{1}{m})^m \leq \frac{1}{e}
\end{equation}
which holds for all $m > 0$. From this inequality, it follows that
\begin{equation}
(1-\frac{1}{n})^t = ((1-\frac{1}{n})^n)^{\frac{t}{n}} \le e^{-\frac{t}{n}}
\end{equation}
With this, the probability $P_{c,t}$ of missing the counterexample with $t$ test runs can be defined as
\begin{equation}
P_{c,t} \le e^{-\frac{t}{n}}
\end{equation}
In order to determine a value for $t$, s.t. $P_{c,t}$ falls below a specified threshold $c$, the inequality is solved for $t$. The resulting inequality then depends on the known parameter $n$ and an arbitrary value for $c$:
\begin{align}
    e^{-\frac{t}{n}} &\leq c && \text{with } 0 \leq c\le 1 \nonumber\\
    t &\geq -n\:\ln(c)
\end{align}
For $c = \frac{1}{e}$ ($\approx 36.79\:\%$) the lower bound of $t$ is exactly $n$. For lower error rates, the number of test runs needs to be greater than $n$. \figref{fig:graph_t_c} shows the lower bounds of $t$ as a function of the probability threshold and the size of the search space. One can see that the increase of test runs is approximately linear for higher thresholds, but transitions to an exponential testing effort to reach lower thresholds.
\begin{figure}[h]
\includegraphics[width=0.95\linewidth]{figures/4-probabilistic_model/graph_t_c}
\caption{Number of required test runs as a function of the probability of false positives $c$ and size of the search space $n$}
\label{fig:graph_t_c}
\end{figure}

\section{Validators vs. test runs}
The last section derived a formula which determines the lower bound of test runs to obtain a specified error rate. However, $t$ is not a parameter that can be adjusted to an arbitrary value on the blockchain. Assuming a blockchain has $m$ validators, the lowest possible bound of test runs is $m$, as every validator invokes the assertion at least once. In cases where $m$ test runs are not sufficient to fall below the threshold $c$, each validator has to invoke the assertion multiple times. More specifically, the number of test runs per validator is given by \eqref{eq:t_validator} and is always a multiple of $m$.
\begin{equation}\label{eq:t_validator}
t_{validator} = \lceil t / m \rceil
\end{equation}

\section{Alternatives to random testing}\label{sec:alt_random}
For applications where low error rates are crucial and random sampling generates too much cost, an alternative could be to coordinate a systematic iteration of the domain. However, the implementation of such a coordination is not trivial, as the validators are engaged via a broadcast and cannot be passed an individual value as input. Instead, possible approaches could be to implement a central instance that allocates an element in the search space to each validator, or to use some unique and intrinsic attribute of validators as a seed or input of a mapping to elements in the search space.

\subsection{Central allocation of elements}
For this approach, the toolchain would need to generate a separate contract that is called by the miners and validators as a proxy. When invoked, this proxy contract allocates the next element of the search space (for instance using a modulo-n counter) and passes this number as an additional parameter to the assertion contract. Furthermore, the proxy contract needs to avoid race conditions in case the assertion mechanism is triggered by several external transactions simultaneously. Another drawback of the proxy contract having to update its internal storage is the consequential serialization of the tests runs.

\subsection{Using unique attributes of a validator}\label{sec:alt_attributes}
This approach strongly depends on whether the validators have a unique id or another attribute, which can be accessed from within a contract. If there is (or the language can be extended with such a feature), the assertion contract needs to implement a non-injective surjective function $f: X_{:A} \rightarrow Y_{:B}$ mapping the respective attribute, represented by a type A, to an element in the search space, represented by type B. Furthermore, in cases where $t > m$, the function should add an offset for each test run, s.t. each iteration checks a distinct element.

>>>>

\chapter{Probabilistic Model}\label{chap:prob_model}
Since the proposed approach implements a probabilistic validation, the reliability of the validation depends on the number of validators in the network in relation to the size of input. Given that the number of validators is finite, the probability of accepting an invalid result may exceed an acceptable threshold for big inputs. Thus, the distributed assertion checking scheme should provide a mechanism to have each validator invoke the assertion contract a specified number of times. To this end, it is necessary to identify a formula which, given the input size and a probability threshold, returns a lower bound for the required number of test runs per validator.

@@

\section{Sample Spaces}\label{sec:sample_space}
The sample space of an assertion is given by the domains of discourse of the quantifiers. \secref{sec:transf_domains} described, that the premises defining the domains are translated into a list of range constraints for the random generators. Based on these constraints, the compiler can derive a formula to determine the size of the sample space at runtime. In simple cases, the sample space is given by a single domain. Such a space is given, for instance, in the introductory example checking whether a number is a prime, i.e., $\mathcal{S} = \lbrace n \in\mathbb{Z} : 2 \le n \le \sqrt{p} \rbrace$. Based on the constraints $2 \le n$ and $n \le \sqrt{p}$, the size of the search space can be determined at runtime with $|\mathcal{S}| = \sqrt{p} - 1$. If the assertion has to sample from more than one domain (for distinct predicate variables), the search space is given by the Cartesian product of these domains. In the example of intersecting sets from \secref{sec:existential}, the sample space is $\mathcal{S} = \mathcal{S}_U \times \mathcal{S}_V$, where $\mathcal{S}_U = \lbrace n \in\mathbb{Z} : 0 \leq n \le |U| \rbrace$ and $\mathcal{S}_V = \lbrace m \in\mathbb{Z} : 0 \leq m \le |V| \rbrace$). At runtime, the size of the search space can be determined with $|\mathcal{S}| = (|U| - 0) * (|V| - 0)$.

@@

\section{Configuration of the Error Rate}\label{sec:prob_threshold}
The goal is to find a number of test runs $t$, s.t. the probability $P_{c,t}$ of accepting an invalid result is below a certain threshold $c$. A validator finds the counterexample in one test run with probability $\frac{1}{n}$, and misses it with probability $(1-\frac{1}{n})$. After $t$ tests runs, the probability that the counterexample has not been found is thus $(1-\frac{1}{n})^t$. Following the same approach as Mahlmann and Schindelhauer \cite{mahl_schindel_2007} to retrieve an upper bound for $P_{c,t}$, the following inequality is used:
\begin{equation}
(1-\frac{1}{m})^m \leq \frac{1}{e}
\end{equation}
which holds for all $m > 0$. From this inequality, it follows that
\begin{equation}
(1-\frac{1}{n})^t = ((1-\frac{1}{n})^n)^{\frac{t}{n}} \le e^{-\frac{t}{n}}
\end{equation}
With this, the probability $P_{c,t}$ of missing the counterexample with $t$ test runs can be defined as
\begin{equation}
P_{c,t} \le e^{-\frac{t}{n}}
\end{equation}
In order to determine a value for $t$, s.t. $P_{c,t}$ falls below a specified threshold $c$, the inequality is solved for $t$. The resulting inequality then depends on the known parameter $n$ and an arbitrary value for $c$:
\begin{align}\label{eq:test_runs}
    e^{-\frac{t}{n}} &\leq c && \text{with } 0 \leq c\le 1 \nonumber\\
    t &\geq -n\:\ln(c)
\end{align}
For $c = \frac{1}{e}$ ($\approx 36.79\:\%$) the lower bound of $t$ is exactly $n$. For lower error rates, the number of test runs needs to be greater than $n$. \figref{fig:graph_t_c} shows the lower bounds of $t$ as a function of the probability threshold and the size of the search space. One can see that the increase of test runs is approximately linear for higher thresholds, but transitions to an exponential testing effort to reach lower thresholds.
\begin{figure}[h]
\includegraphics[width=0.95\linewidth]{figures/4-probabilistic_model/graph_t_c}
\caption{Number of required test runs as a function of the probability of false positives $c$ and size of the search space $n$}
\label{fig:graph_t_c}
\end{figure}

\section{Validators vs. Test Runs}
The last section derived a formula which determines the lower bound of test runs to obtain a specified error rate. However, $t$ is not a parameter that can be adjusted to an arbitrary value on the blockchain. Assuming a blockchain has $m$ validators, the lowest possible bound of test runs is $m$, as every validator invokes the assertion at least once. In cases where $m$ test runs are not sufficient to fall below the threshold $c$, each validator has to invoke the assertion multiple times. More specifically, the number of test runs per validator is given by \eqref{eq:t_validator} and is always a multiple of $m$.
\begin{equation}\label{eq:t_validator}
t_{validator} = \lceil t / m \rceil
\end{equation}

\section{Alternatives to Random Testing}\label{sec:alt_random}
For applications where low error rates are crucial and random sampling generates too much cost, an alternative could be to coordinate a systematic iteration of the domain. However, the implementation of such a coordination is not trivial, as the validators are engaged via a broadcast and cannot be passed an individual value as input. Instead, possible approaches involve the implementation of a central instance that allocates an element in the search space to each validator, or to use some unique and intrinsic attribute of validators as a seed or input of a mapping to elements in the search space.

\subsection{Central Allocation of Elements}
When using a central instance for the allocation of elements, the toolchain would need to generate a separate contract that is called by the bakers and validators as a proxy. When invoked, this proxy contract allocates the next element of the search space (for instance using a modulo-n counter) and passes this number as an additional parameter to the assertion contract. Furthermore, the proxy contract needs to avoid race conditions in case the assertion mechanism is triggered by several external transactions simultaneously. Another drawback of the proxy contract having to update its internal storage is the consequential serialization of the tests runs.

\subsection{Using Unique Seeds}\label{sec:alt_attributes}
Instead of installing a central instance, validators could use an intrinsic unique ID or attributed as a seed for a random generation or as an input for a mapping. This approach strongly depends on whether the validators have such an ID or attribute, which is accessible from within the VM. In the case of a mapping, the assertion contract needs to implement a non-injective surjective function $f: X_{:A} \rightarrow Y_{:B}$ mapping the respective attribute, represented by a type A, to an element in the search space, represented by type B. Furthermore, in cases where $t > m$, the function should add an offset for each test run, s.t. each iteration checks a distinct element.
